Future challenges for Research Software Engineering

### 2025-04-17 future_challenges_for_RSEng.tex
add some sentences to ethics section (#12)
* add some sentences to ethics section

Co-authored-by: Gerasimos Chourdakis <gerasimos.chourdakis@ipvs.uni-stuttgart.de>

* some color and structuring boxes

* start

---------

Co-authored-by: Frank Löffler <frank.loeffler@uni-jena.de>
Co-authored-by: Gerasimos Chourdakis <gerasimos.chourdakis@ipvs.uni-stuttgart.de>
Co-authored-by: Florian Goth <florian.goth@physik.uni-wuerzburg.de>
Co-authored-by: Florian Goth <florian.goth@uni-wuerzburg.de>

\newtcbtheorem[number within=section]{whatis}{What is 2035}{
  enhanced,
  segmentation engine=empty,
%  beamer,
fuzzy shadow={2mm}{-1mm}{0mm}{0.1mm}%
{black!50!white},
  sharp corners,
  attach boxed title to top left={
    yshifttext=-1mm
  },
%  halign lower=flush center,
  colback=white,
  colframe=Maroon!75!white,
  fonttitle=\bfseries,
  attach boxed title to top left={yshift=-2mm,xshift=3mm},
  fontlower=\itshape,
  boxed title style={
    sharp corners,
    size=small,
    colback=Maroon!75!black,
    colframe=Maroon!75!black,
  } 
}{ist}

\newtcbtheorem[number within=section]{story}{Story}{
  enhanced,
  segmentation engine=empty,
%  beamer,
fuzzy shadow={2mm}{-1mm}{0mm}{0.1mm}%
{black!50!white},
  sharp corners,
  attach boxed title to top right={
    yshifttext=-1mm
  },
%  halign lower=flush center,
  colback=white,
  colframe=ForestGreen!75!white,
  fonttitle=\bfseries,
  attach boxed title to top right={yshift=-2mm,xshift=3mm},
  fontlower=\itshape,
  boxed title style={
    sharp corners,
    size=small,
    colback=ForestGreen!75!black,
    colframe=ForestGreen!75!black,
  }
}{str}

@@

\begin{story}{The greater world of Kim and Kay}{ethics}
10 years have passed since Kim and Kay have started on their journey of being RSEs. Back then,
at the universities of Eden and Utopia, life was more orderly and harmonious. Kay is working in a project that develops
software for cheap water monitoring devices in the global south. Kim is currently designing the software service stack
at her university, and increasingly struggles to find compelling Open Source alternatives that let's them keep their data
to software that is backed by large corporate entities. So far she still has the support of her university board, since it values
strategic autonomy.
Her son Kyle on the other hand started out with a degree in Physics and then took one of the newly created RSE Master courses,
but quickly decided that work in academia with its time limited contracts is not attractive.
Governments around the world are struggling to reposition themselves in this emerging deglobalized world with its new blocks
and therefore have deprioritized publicly funded research. Kyle decided to go to a defence company since they offer a
more secure employment in this environment.
\end{story}

\subsection{What are new tasks of RTPs?}
 FG-NOTE: RSEs are one type of specialization of RTPs in science. Do we see any other?

\subsection{Impact of Alternate Career Paths for Researchers on the Structure of Research Institutes}

\subsection{How have academic Institutions transformed by then? ( Point 4 in the issue )}
\begin{whatis}{What do we want to have achieved}{acedemic}
We hope to see that universities use the presence of an RSE Department actively as a means to differentiate themselves from other universities.
We hope that the central RSE units have local RSE representatives, or even small RSE groups that handle tasks within a department to oversee specific projects[FIXME: cite group paper].
We also hope, that specialized clusters form, with certain RSE units specializing in certain tasks like data management, HPC, or visualization.
At universities which have them, we want to see, that researchers naturally accept them as part of their used services.
\end{whatis}

>>>>

\newcounter{whatiscounter}
\newtcbtheorem[use counter=whatiscounter, number format=\Alph]{whatis}{What is 2035}{
  enhanced,
  segmentation engine=empty,
%  beamer,
fuzzy shadow={2mm}{-1mm}{0mm}{0.1mm}%
{black!50!white},
  sharp corners,
  attach boxed title to top left={
    yshifttext=-1mm
  },
%  halign lower=flush center,
  colback=white,
  colframe=Maroon!75!white,
  fonttitle=\bfseries,
  attach boxed title to top left={yshift=-2mm,xshift=3mm},
  fontlower=\itshape,
  boxed title style={
    sharp corners,
    size=small,
    colback=Maroon!75!black,
    colframe=Maroon!75!black,
  } 
}{ist}

\newcounter{storycount}
\newtcbtheorem[use counter=storycount]{story}{Story}{
  enhanced,
  segmentation engine=empty,
%  beamer,
fuzzy shadow={2mm}{-1mm}{0mm}{0.1mm}%
{black!50!white},
  sharp corners,
  attach boxed title to top right={
    yshifttext=-1mm
  },
%  halign lower=flush center,
  colback=white,
  colframe=ForestGreen!75!white,
  fonttitle=\bfseries,
  attach boxed title to top right={yshift=-2mm,xshift=3mm},
  fontlower=\itshape,
  boxed title style={
    sharp corners,
    size=small,
    colback=ForestGreen!75!black,
    colframe=ForestGreen!75!black,
  }
}{str}

@@

\begin{story}{The greater world of Kim and Kay}{ethics}
10 years have passed since Kim and Kay have started on their journey of being RSEs. Back then,
at the universities of Eden and Utopia, life was more orderly and harmonious. Kay is working in a project that develops
software for cheap water monitoring devices in the global south. Kim is currently designing the software service stack
at her university, and increasingly struggles to find compelling Open Source alternatives that let's them keep their data
to software that is backed by large corporate entities. So far she still has the support of her university board, since it values
strategic autonomy.
Her son Kyle on the other hand started out with a degree in Physics and then took one of the newly created RSE Master courses,
but quickly decided that work in academia with its time limited contracts is not attractive.
Governments around the world are struggling to reposition themselves in this emerging deglobalized world with its new blocks
and therefore have deprioritized publicly funded research. Kyle decided to go to a defence company since they offer a
more secure employment in this polarized world.
\end{story}

\subsection{How have academic Institutions transformed by then?}
\begin{whatis}{What do we want to have achieved}{acedemic}
We hope to see that universities use the presence of an RSE Department actively as a means to differentiate themselves from other universities.
We hope that the central RSE units have local RSE representatives, or even small RSE groups that handle tasks within a department to oversee specific projects[FIXME: cite group paper].
We also hope, that specialized clusters form, with certain RSE units specializing in certain tasks like data management, HPC, or visualization.
At universities which have them, we want to see, that researchers naturally accept them as part of their used services.
\end{whatis}
